Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-22 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
more text to appendix

For the nonrotating QNMs we used the numerical data of Ref.~\cite{Blazquez-Salcedo:2016enn} and
generated a new quartic fit in $\gamma_{\rm EdGB} = \ell_{\rm EdCS} / M$ using numerical QNM data valid for small
values of the coupling constant [see, in particular, Eq.~(27) and Fig.~2 of Ref.~\cite{Blazquez-Salcedo:2016enn}].
%
We find,
%
\begin{subequations}
\begin{align}
    M {\rm Re}(\sigma)_{\rm EdGB} &= M {\rm Re}(\sigma)_{\rm GR} \, \left( 1 + 0.0107 \, \gamma_{\rm EdGB}^4 \right),
    \\
    M {\rm Im}(\sigma)_{\rm EdGB} &= M {\rm Im}(\sigma)_{\rm GR} \, \left( 1 - 0.0044 \, \gamma_{\rm EdGB}^4 \right).
\end{align}
\end{subequations}
%
The small values of the numerical prefactors of $\gamma_{\rm EdGB}^4$ are a consequence of the how weakly the QNMs
of BHs in EdGB gravity deviate from their GR counterparts, even at moderate values $\gamma_{\rm EdGB} \approx 0.3$.

The rotating corrections are included using the results of Ref.~\cite{Pierini:2021jxd}, in particular their Eqs.~(51) and~(52),
which for consistency with our previous discussion, we truncated at quartic-order in $\gamma_{\rm EdGB}$.
%
The resulting formula can then be expanded in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}.
%
These steps yield,
%
\begin{align}
    p_{\rm EdGB} = 4, \quad \delta \omega_{\rm EdGB}^{(0)} = -0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
    \nonumber \\
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.101.
\end{align}

%
% Taken together, use as our fitting expressions for the QNM is EdGB gravity
% %
% \begin{subequations}
% \begin{align}
%     M {\rm Re}(\sigma)_{\rm EdGB} &= {\rm Re}(\omega)_{\rm GR} \, \left[ 1 + 0.0107 (\ell_{\rm EdGB} / M)^4 \right],
%     \\
%     M {\rm Im}(\sigma)_{\rm EdGB} &= \dots % {\rm Im}(\omega)_{\rm GR} \, \left[ 1 - 0.0044 (\ell_{\rm EdGB} / M)^4 \right].
% \end{align}
% \end{subequations}
%
% \dots

@@

We now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\gamma_{\rm dCS}$, and
gather the terms proportional to $\gamma_{\rm dCS}$.
%
We obtain
%
\begin{align}
    M \, \omega_{\rm dCS} =
    \left( 0.3722 + 1.1945 \gamma_{\rm dCS}^4 \right)
    + \left(0.1861 + 5.1828 \gamma_{\rm dCS}^4 \right) \, \chi,
    \nonumber \\
    \label{eq:omega_dcs_interm}
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$.
%
We find that (reassuringly) that the nonrotating GR part of the expression above agree with $\omega^{(0)}$ of Ref.~\cite{Maselli:2019mjd}
to 0.5\% relative percent error. The same error estimate is larger ($\approx 20$\%) for the linear-in-spin coefficient;
0.1861 in comparison to 0.1258 of Ref.~\cite{Maselli:2019mjd}.
%
We attribute this difference to Ref.~\cite{Wagle:2021tam} having fitted Eq.~\eqref{eq:omega_dsc_eg} to QNM data compute to linear-order in spin, whereas~\cite{Maselli:2019mjd}
fitted Eq.~\eqref{eq:kerr_expansion} to Kerr QNM valid to all orders in spin.

We can now isolate the dCS corrections from Eq.~\eqref{eq:omega_dcs_interm} and
compare against Eq.~\eqref{eq:qnm_nongr_iso}, to find
%
\begin{equation}
p_{\rm dCS} = 4\,, \quad \delta \omega^{(0)}_{\rm dCS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm dCS} = 41.199.
\label{eq:cs_omega_coefs}
\end{equation}
%
We can carry the same steps for $\tau_{\rm dCS}$ and find
%
\begin{equation}
\delta \tau^{(0)}_{\rm dCS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm dCS} = 794.66,
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed non-GR parameters in the ringdown of the \pSEOB{} waveform model for this theory.
%
As for the case of EdGB gravity, we remark that the alarmingly large values of
$\delta \omega_{\rm dCS}^{(1)}$ and $\delta \tau_{\rm dCS}^{(1)}$ are
compensated by the assumptions that $(\ell_{\rm dCS} / M_{\rm f})^2$ and $\chi$
are much less than unity, which are the assumptions used in Ref.~\cite{Wagle:2021tam}
to compute the QNMs.

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), particularizing to $\lame = \lamo = 1$.
%
We further linearize that equation in $\chi$ and particularize $m=2$.
%
This process, shows that the mode labelled ``-'' is the least damped and it is the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in,
%
\begin{align}
    p_{\rm cEFT} = 4, \quad \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12.
\end{align}

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we found that both modes (``-'' and ``+'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast and we chose the ``-'' mode which decreases less the damping time.
%
We then find
%
\begin{align}
    p_{\rm qEFT} = 6, \quad \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}

>>>>

For the nonrotating QNMs we used the numerical data of Ref.~\cite{Blazquez-Salcedo:2016enn} and
generated a new quartic fit in $\gamma_{\rm EdGB} = \ell_{\rm EdCS} / M$ using numerical QNM data valid for small
values of the coupling constant [see, in particular, Eq.~(27) and Fig.~2 of Ref.~\cite{Blazquez-Salcedo:2016enn}].
%
We find,
%
\begin{subequations}
\begin{align}
    M {\rm Re}(\sigma)_{\rm EdGB} &= M {\rm Re}(\sigma)_{\rm GR} \, \left( 1 + 0.0107 \, \gamma_{\rm EdGB} \right),
    \\
    M {\rm Im}(\sigma)_{\rm EdGB} &= M {\rm Im}(\sigma)_{\rm GR} \, \left( 1 - 0.0044 \, \gamma_{\rm EdGB} \right).
\end{align}
\end{subequations}
%
The small values of the numerical prefactors of $\gamma_{\rm EdGB}^4$ are a consequence of the how weakly the QNMs
of BHs in EdGB gravity deviate from their GR counterparts, even at moderate values $\gamma_{\rm EdGB} \approx 0.3$.

The rotating corrections are included using the results of Ref.~\cite{Pierini:2021jxd}, in particular their Eqs.~(51) and~(52),
which for consistency with our previous discussion, we truncated at quartic-order in $\gamma_{\rm EdGB}$.
%
The resulting formula can then be expanded in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}.
%
These steps yield $p_{\rm EdGB} = 4$,
%
\begin{align}
     \delta \omega_{\rm EdGB}^{(0)} = -0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
\end{align}
%
for the $j=0$ coefficients and
%
\begin{equation}
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.101.
\end{equation}
%
for the $j=1$ coefficients.

% Taken together, use as our fitting expressions for the QNM is EdGB gravity
% %
% \begin{subequations}
% \begin{align}
%     M {\rm Re}(\sigma)_{\rm EdGB} &= {\rm Re}(\omega)_{\rm GR} \, \left[ 1 + 0.0107 (\ell_{\rm EdGB} / M)^4 \right],
%     \\
%     M {\rm Im}(\sigma)_{\rm EdGB} &= \dots % {\rm Im}(\omega)_{\rm GR} \, \left[ 1 - 0.0044 (\ell_{\rm EdGB} / M)^4 \right].
% \end{align}
% \end{subequations}
%
% \dots

@@

We now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\gamma_{\rm dCS}$, and
gather the terms proportional to $\gamma_{\rm dCS}$.
%
We obtain
%
\begin{align}
    M \, \omega_{\rm dCS} =
    \left( 0.3722 + 1.1945 \gamma_{\rm dCS} \right)
    + \left(0.1861 + 5.1828 \gamma_{\rm dCS} \right) \, \chi,
    \nonumber \\
    \label{eq:omega_dcs_interm}
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$.
%
We find that (reassuringly) that the nonrotating GR part of the expression above agree with $\omega^{(0)}$ of Ref.~\cite{Maselli:2019mjd}
to 0.5\% relative percent error. The same error estimate is larger ($\approx 20$\%) for the linear-in-spin coefficient;
0.1861 in comparison to 0.1258 of Ref.~\cite{Maselli:2019mjd}.
%
We attribute this difference to Ref.~\cite{Wagle:2021tam} having fitted Eq.~\eqref{eq:omega_dsc_eg} to QNM data compute to linear-order in spin, whereas~\cite{Maselli:2019mjd}
fitted Eq.~\eqref{eq:kerr_expansion} to Kerr QNM valid to all orders in spin.

We can now isolate the dCS corrections from Eq.~\eqref{eq:omega_dcs_interm} and
compare against Eq.~\eqref{eq:qnm_nongr_iso}, to find
$p_{\rm dCS} = 4$,
%
\begin{equation}
\delta \omega^{(0)}_{\rm dCS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm dCS} = 41.199.
\label{eq:cs_omega_coefs}
\end{equation}
%
We can carry the same steps for $\tau_{\rm dCS}$ and find
%
\begin{equation}
\delta \tau^{(0)}_{\rm dCS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm dCS} = 794.66,
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed non-GR parameters in the ringdown of the \pSEOB{} waveform model for this theory.
%
As for the case of EdGB gravity, we remark that the alarmingly large values of
$\delta \omega_{\rm dCS}^{(1)}$ and $\delta \tau_{\rm dCS}^{(1)}$ are
compensated by the assumptions that $\gamma_{dCS}$ and $\chi$
are much less than unity, which are the assumptions used in Ref.~\cite{Wagle:2021tam}
to compute the QNMs.

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), particularizing to $\lame = \lamo = 1$.
%
We further linearize that equation in $\chi$ and particularize $m=2$.
%
This process, shows that the mode labeled ``-'' is the least damped and it is the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12.
\end{align}

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we found that both modes (``-'' and ``+'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast and we chose the ``-'' mode which decreases less the damping time.
%
We then find this time that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}
%
As with the previous theories, the larger values of some of these parameters are compensated
by the assumptions of small coupling and small spin used to calculate the QNM frequencies.
